\section{An analysis of supplemental CIs}

One of the most prominent accounts of presuppositions is the one by
\citet{sandt1992presupposition}, who treats presupposition projection as
anaphora resolution.
%%explain more.

CIs share their projection behaviour with other projection phenomena, so this
requires a unified analysis of their behaviour.

%\citet{potts2005logic} describes the following properties of CIs:
%
%\begin{itemize}
%  \item CIs tend to project globally (scopelessness)
%  \item CIs are non-cancellable
%  \item CIs are non-restrictive
%  \item Speaker-orientedness
% \end{itemize}

We will argue that all these properties can be described in terms of
a species of indexicality \citep[cf.][]{amaral2007review}, which in the case
of supplemental CIs comes down to indexicality to the projection site of the
anchor.

\subsection{Projection as anchoring}

\subsubsection{CIs are `piggybacking' on their anchor}

* CIs always occur with an anchor\\
* Similarity to E-type pronouns\\
* CIs always project as far as their anchor
%* CIs provide novel information to the context of their anchor


\subsubsection{CIs require their anchor to project}

Infelicitous with non-specific indefinites, quantificational constructions

The information structure of CIs (i.e., their subordinatedness) signals
projection; this is why we can distinguish between RRCs and NRRCs --- by
means of the comma in this case. Non-restrictiveness requires projection.
This is because a non-projecting anchor requires the introduction of a new
entity, which by definition lacks a complete description, so any new
information will contribute to the description of the referent and will
therefore be restrictive. On the other hand, in case the anchor projects,
this means that the referent has been established, in the sense that it can
be identified in the common ground, so providing new information does not
affect the choice of referent, so it works non-restrictively.

\ex. example with indefinite anchor -> only specific reading

\ex. example with pronominal anchor -> only felicitous if referent can be
  determined unambiguously.

Because of the requirement for a specific anchor, CIs are often attached to
presuppositional anchors, which explains their tendency to project to the
global discourse context. However, there are also examples of CIs that do
not project to the global discourse level, as in \Next \citep[example
from][]{amaral2007review}.

\ex. Joan is crazy. She's hallucinating that some geniuses in Silicon Valley
have invented a new brain chip that's been installed in her left temporal
lobe [\ldots]. Joan believes that her chip\underline{, which she has
installed last month,} has a twelve year guarantee.

In this example, the CI triggered by the underlined NRRC does not project to
the global discourse context, since it is clear that the speaker does not
want to convey that Joan actually had the chip installed last month. Note,
however, that the direct anchor of the CI content (``her chip''), is
a possessive construction --- a presupposition that in the current context
binds to the introduction of ``a new brain chip'' in the sentence before
\citep[following][]{sandt1992presupposition}.  Thus, the anchor is not
accommodated to the global discourse level, and therefore the CI attached to
the anchor isn't either. Yet, the fact that the anchor projects allows it to
felicitously occur with a (restrictive) CI.

%Something about generics? and inclusion and attribution as a contrast to
%identification with anchor? \cite[see][]{heringa2012appositional}

\subsection{Formalizing projection-anaphoricity}

\subsubsection{Projective Discourse Representation Theory}

[short introduction]

\subsubsection{Enhancing PDRSs with information structure}

* Adding MAPs to PDRSs. \\
* Projection as \textit{strict} subordination


\subsection{Representing the projection behaviour of CIs}

\subsubsection{Supplemental clauses}

\paragraph{Appositives.}~

\noindent \parbox[b]{\textwidth}{
\ex. Sheila believes that Chuck\underline{, a psychopath,} is fit to watch
the kids.\\
\pdrs{1}{$3\gets$x~~$1\gets$p}{$3\gets$Sheila(x)\\ $1\gets$believes(x,p)\\
$1\gets$p:\pdrs{2}{$4\gets$y~~$5\gets$z}{$4\gets$Chuck(y)\\ 
$5\gets$psychopath(z)\\ $5\gets$z=y\\ $2\gets$fit-to-watch-kids(y)}{$2<4$
~~$2<5$~~$5=4$}}{$1< 3$}

}

\paragraph{NRRCs.}~

\noindent\parbox[b]{\textwidth}{
\ex. Ames\underline{, who stole from the FBI,} is behind bars.\\
\pdrs{1}{$3\gets$x~~$4\gets$y}{$3\gets$Ames(x)\\ $4\gets$FBI(y)\\ 
$5\gets$stole\_from(x,y)\\ $1\gets$behind\_bars(x)}{$1<3$~~$1<5$~~$5<4$
~~$5=3$}

}

\paragraph{\textit{As}-parentheticals.}~

\noindent\parbox[b]{\textwidth}{
\ex. It is not the case that Ames\underline{, as the press reported,} is
a spy.\\
\textit{cf.} Ames, of whom the press reported he is a spy, is not (a spy).\\
\pdrs{1}{}{$1\gets\neg$\pdrs{2}{$3\gets$x~~$4\gets$y~~$5\gets$p}{
$3\gets$Ames(x)\\ $4\gets$press(y)\\ $5\gets$p:\pdrs{6}{}{$6\gets$spy(x)}{
$6<3$}\\ $5\gets$reported(y,p)\\ $2\gets$spy(x)}{$2<3$~~$2<5$~~$5<4$
~~$5=3$}}{}

}
%[..]
%
%
%Note that \citet{potts2005logic} also introduces a \textsc{comma} operator,
%although he uses it to add supplemented content to the CI dimension.
%\citet{amaral2007review} argue against the use of a \textsc{comma} operator
%by adhering to both its overinterpretation (not all comma's are indicators
%of CIs -- not even all comma's that introduce a subordinate clause to
%a noun), and underinterpretation (since not all CIs are triggered by
%a comma \citep[cf. the `malheureusement' example by][]{bonami2004adverb},
%and because of type-matching).
%
%
\subsubsection{Supplemental adverbs}

\paragraph{Topic/Speaker-oriented Adverbs.}


\paragraph{Utterance level modifiers.} (e.g., ``Frankly, I don't like him'',
``In case you're hungry, I'm going to the grocery.'', etc.) Following
\citet[pp.725-729]{amaral2007review}, these modifiers behave differently
than other CIs, both on the theoretical and the implementational level.


%
